Virtual Memory Assisted Transactional Memory for Dynamic Languages

### 2014-05-27 talk/dls2014/paper/paper.tex
again change the metric for explaining memory usage (not sure what's best)

\begin{figure*}[t]
  \begin{centering}
    \includegraphics[scale=0.8]{\string"segment addressing\string".pdf}
    \par\end{centering}

    \protect\caption{Segment Addressing\label{fig:Segment-Addressing}}
\end{figure*}

@@

\begin{figure}[h]
  \begin{centering}
    \includegraphics[scale=0.8]{\string"mmap pages\string".pdf}
    \par\end{centering}

    \protect\caption{\texttt{mmap()} Page Mapping\label{fig:mmap()-Page-Mapping}}
\end{figure}

@@

\begin{figure}[h]
  \begin{centering}
    \includegraphics[width=1\columnwidth]{\string"page remapping\string".pdf}
    \par\end{centering}

    \protect\caption{Page Remapping: (I) after \texttt{mmap()}. (II) remap all pages to
      segment~0, fully shared memory configuration. (III) privatise single
      pages.\label{fig:Page-Remapping}}
\end{figure}

@@

\begin{figure*}[t]
  \begin{centering}
    \includegraphics[scale=0.8]{\string"segment layout\string".pdf}
    \par\end{centering}

    \protect\caption{Segment Layout\label{fig:Segment-Layout}}
\end{figure*}

@@

Third, since we do copy on write, if all objects were to be
written to at the same time, all pages would need to be privatised
for all objects in the old object space. In that case we would
need the total amount of memory required by old objects multiplied
by $N+1$ (incl. the sharing segment). Pages get re-shared during
major collections if possible.

In figure \ref{fig:richards_mem} we look at the memory usage of one of
our benchmarks called Richards\footnote{OS kernel simulation
benchmark}. The \emph{GC managed memory} counts all memory used in the
old object space including the memory required for private pages. The
sharp drops in memory usage come from major collections that free old
objects and re-share pages. The average memory usage stays around
29~MiB and we see that in this benchmark we have around 1 major
collection every second. The \emph{page privatisation}, which
represents the percentage of used pages with at least one private
copy, shows the same spikes as the memory usage. These come directly from
re-sharing the pages. The maximum page privatisation is around $20\%$
between major collections. Thus we can say that $~20\%$ of the old
objects get modified between collections in this benchmark.

>>>>

\begin{figure*}[t]
  \centering
  \includegraphics[scale=0.8]{\string"segment addressing\string".pdf}
  \caption{Segment Addressing\label{fig:Segment-Addressing}}
\end{figure*}

@@

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.8]{\string"mmap pages\string".pdf}
  \caption{\texttt{mmap()} Page Mapping\label{fig:mmap()-Page-Mapping}}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=1\columnwidth]{\string"page remapping\string".pdf}
  \caption{Page Remapping: (I) after \texttt{mmap()}. (II) remap all pages to
    segment~0, fully shared memory configuration. (III) privatise single
    pages.\label{fig:Page-Remapping}}
\end{figure}

@@

\begin{figure*}[t]
  \centering
  \includegraphics[scale=0.8]{\string"segment layout\string".pdf}
  \caption{Segment Layout\label{fig:Segment-Layout}}
\end{figure*}

@@

Third, since we do copy on write, if all objects were to be written to
at the same time, all pages would need to be privatised for all
objects in the old object space. In that case we would need the total
amount of memory required by old objects multiplied by $N+1$
(incl. the sharing segment). During major collections we re-share the
pages if possible.

In figure \ref{fig:richards_mem} we look at the memory usage of one of
our benchmarks called Richards\footnote{OS kernel simulation
benchmark}. The \emph{GC managed memory} counts all memory used in the
old object space including the memory required for private pages. The
sharp drops in memory usage come from major collections that free old
objects and re-share pages. The average memory usage stays around
29~MiB and we see that in this benchmark we have around 1 major
collection every second.

The \emph{page privatisation}, which represents the ratio between
private pages and shared pages, gives us an idea about how many
private copies exist for each shared page. If the ratio is $>1.0$, it
means that on average we have more than one private copy for each
shared page. In the worst case described above, this number would
reach $N$ and means we need $N\times$ the memory for the private
pages alone. In this benchmark, we see the same spikes as the memory
usage. These come directly from re-sharing the pages. The maximum page
privatisation is around $0.5$ between major collections, thus the
private pages are responsible for a $50\%$ increase in the required
memory. Since the spikes in the GC managed memory line actually
show increases by $~80\%$, it means that the rest comes from actual
garbage objects that were collected.
